Effects of the body force on the pedestrian and the evacuation dynamics

### 2020-03-05 paper.tex
correcciones chicas de ortografia

\item[(c)] We will explore crowd densities allowing body compressions up to
$0.1\,$m. This corresponds to remarkably high densities, since our own  (real-
life) estimates do not surpass $0.065\,$m (see  Section~\ref{experimental}).
Therefore, the extremely high density  scenarios (say, above $0.065\,$m) will be
considered for the purpose of a  tendency, but not as a common real-life
situation. No casualties due to high  pressures will be further considered.

@@

We stress the fact that our concern is placed on the collective dynamics of
crowded environments. We will work on the hypothesis that any ``reasonable''
parameter set should reproduce the collective behavior (say, the slowing down
in the fundamental diagram; see Section~\ref{corridor}). We will not attempt to
optimize parameters from other objective functions. \\

We will further sustain the hypothesis of soft matter in the context of crowded
environments. The body compression and the sliding friction parameters are
supposed to be connected in some way under this hypothesis. However, we will
not introduce any direct link between $\mathcal{K}$ and $\mathcal{K}_c$. We
will investigate the interplay between both in Section~\ref{corridor}.  \\

@@

Fig.~\ref{vd_vs_te}  shows either FIS or FIF, and a FIS+FIF pattern for desired
velocities $v_d>1.2\,$m/s. The  evacuation time attains a FIS+FIF pattern for
compression coefficients below  $k=1.2\,$E5. This means that ``soft''
individuals can attain this  behavioral pattern.  Notice that higher values of
$k_n$ allow  only FIS or FIF patterns. For the highest explored value
$k_n=1.2\,$E6, no  FIS can be seen at all. Besides, the evacuation pattern for
$k_n = 0$ and $k_n  = 2.6\,$E4 are very similar since the body force intensity
is of the same order or less than the social  force for these stiffness values.
\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the
pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the
pedestrians desired velocity. Each symbol indicates the $k_n$ value
corresponding to the body force (see the labels). The data corresponds to a
bottleneck with periodic boundary conditions (re-injecting pedestrians). The
average was taken over time and the pedestrians in the bottleneck. The sampling
was done every five seconds once the crowd reached the stationary state (say,
$t=20\,$s) until the end of the simulation ($t=1000\,$s). Color online only.}
\label{degree_overlap_vd}
\end{figure}

@@

We first computed the contact network in the same way  as in the bottleneck
geometry. Fig.~\ref{degree_dens} shows the  mean degree as a function of the
global density for different $k_n$ values (see  caption for details). The mean
degree vanishes at very low  densities because the pedestrians do not touch each
other. When the density  surpasses 4.5, a few pedestrians start to touch each
other,  raising the mean degree. As the density continuous increasing,  the mean
degree approaches the asymptotic value  of six. Degree six corresponds to the
maximum packing density for identical  hard circles.\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the 
corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that 
connect the nodes (pedestrians) represent the contact between them. The colors 
stand for the degree of the node (the number of pedestrians that are in contact 
with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary 
conditions and $v_d=1$. (a) corresponds to a simulation without body force and 
(b) corresponds to a simulation with $k_n=1.2$~E5. The friction coefficient and 
the other SFM parameters are the same as in Section \ref{bottleneck}. Color 
online only.}
\label{network_corridor}
\end{figure}

@@

Fig.~\ref{network_bottleneck} and  Fig.~\ref{network_corridor} illustrate the
connectivity differences between  the bottleneck and the corridor situation. The
bulk in  Fig.~\ref{network_bottleneck} appears more heavily connected among
``soft''  pedestrians than among ``stiff'' pedestrians (see both snapshots at
$v_d=10$).  The opposite occurs in Fig.~\ref{network_corridor}. This discrepancy
seems to  be related to the boundary conditions, since the same SFM parameters
were  applied on both situations. We may speculate that this phenomenon occurs
because, in the corridor, the lateral walls act like a  confining barrier that
forces the ``stiff'' pedestrian to increase his (her) contacts. On the contrary,
no real containment walls exist in the bottleneck situation (regardless the side
walls). \\

We can test our hypothesis by computing the mean overlap. 
Fig.~\ref{overlap_dens} shows this magnitude for the corridor situation. Notice 
that ``soft'' pedestrians attain more overlap than the ``stiff'' 
ones, as expected. This is in agreement with Fig.~\ref{overlap_vd} for the 
bottleneck situation (at a fixed value of $v_d$). The curves in 
Fig.~\ref{overlap_vd}, however, do not meet each other as in  
Fig.~\ref{overlap_dens}. All the curves meet each other at
high-density values. This phenomenon occurs due to space limitations
in the corridor that produces overlapping (almost) independently of 
the value of $k_n$. \\

@@

The above results represent an important step in the  investigation. The inter-
pedestrian connectivity differences between the bottleneck geometry  and the
corridor geometry were not expected and opens two major questions:  how do the
pedestrians interact with the walls as a function of the body stiffness, and
consequently, how does  this affect the flux across the corridor.\\

@@

The $\langle v_x\rangle$ values in Fig.~\ref{kn_vs_vx_corridor}  were computed
at densities $\rho\geq 5.5$. According to  Fig.~\ref{overlap_dens}, at these
densities, the overlaps among pedestrians attain significant values.
Furthermore,  stiffness values $k_n\geq 10^4$ move the system to a more heavily
connected  stage, and thus, pedestrians have no choice but to walk at  (almost)
the same speed as his (her) neighbors. This  behavior may be envisaged as the
passage from a ``free''  walking movement to a  constrained movement as the
stiffness increases. A more  physical picture would assimilate the former as a
``fluid-like  state'' and the latter as a ``solid-like state''. When the
stiffness is  very high (say $k_n=$1.2~E6), all the pedestrians  are expected to
walk at a common velocity. The  pedestrians that walk in physical contact with
the wall are the ones who  determine the velocity of the whole crowd, as
discussed  below.\\

@@

From the results displayed above, we realize that the crowd behaves like a solid
for very stiff pedestrians. This means that the crowd can not be easily
``deformed''. In this context, deformation means that parts of the crowd may be
allowed to move faster than other parts of the crowd. \\

@@

Where $\langle \cdot \rangle$ means the average taken over time. This definition
compares the velocity of the pedestrians close to the wall (boundary) with
respect to the velocity of the pedestrians at the center of the corridor
(center). Thus, $\dot{\gamma}$ vanishes (no deformation) as the stiffness level
increases. This phenomenon is shown in Fig.~\ref{strain_rate_vs_kn} where we can
see that the strain rate drops for high values of $k_n$. \\

We conclude this Section by stressing, once again, that stiffer pedestrians
attain opposite results in corridors with respect to bottlenecks. The walls in
the corridors play a critical role that prevents pedestrians from detaching from
each other. This effect has can be observed as an increment in the connectivity
of the contact network and the flattening of the velocity profile (thus reducing
the strain rate). Very high enough stiffness values stuck the pedestrians
leading to a ``solidification'' of the crowd. All the pedestrians walk at almost
the same velocity at this stage. The pedestrians that are in contact with the
walls are the ones that determine the velocity of the whole crowd. We already
analyzed this situation in Ref~\cite{dorso_2019}. \\

@@

A further increment of $\mathcal{K}$ to $\mathcal{K}=$137 (ten times the 
original SFM value) attains a plateau ($\rho>5$ p m$^{-2}$) before vanishing at 
very high densities. This occurs on either $\mathcal{K}_c=$ 0 and 
$\mathcal{K}_c = $ 68.\\

@@

Having adopted as fundamental philosophy of work that the parameters of the SFM
should be properly determined by requiring that the model properly reproduces
well established experimental data, we explored the effect on the pedestrians
dynamics of the (sometimes neglected) body force in the framework of the SFM. We
showed that the stiffness coefficient ($k_n$) has a significant impact on the
evacuation dynamics (bottleneck) and also in the dynamics of pedestrians walking
along a straight corridor.\\

@@

The authors want to thanks the degree students Josefina Catoni and Ayelen Santos
 for providing data acquired at the subway in Buenos Aires, Argentina. \\

>>>>

\item[(c)] We will explore crowd densities allowing body compressions (overlaps)
up to $0.1\,$m. This corresponds to remarkably high densities, since our own
(real- life) estimates do not surpass $0.065\,$m (see
Section~\ref{experimental}). Therefore, the extremely high density  scenarios
(say, above $0.065\,$m) will be considered for the purpose of a  tendency, but
not as a common real-life situation. No casualties due to high  pressures will
be further considered.

@@

We stress the fact that our concern is placed on the collective dynamics of
crowded environments. We will work on the hypothesis that any ``reasonable''
parameter set should reproduce the collective behavior (say, the slowing down in
the fundamental diagram; see Section~\ref{Dimensionless}). We will not attempt
to optimize parameters from other objective functions. \\

We will further sustain the hypothesis of soft matter in the context of crowded
environments. The body compression and the sliding friction parameters are
supposed to be connected in some way under this hypothesis. However, we will not
introduce any direct link between $\mathcal{K}$ and $\mathcal{K}_c$. We will
investigate the interplay between both in Section~\ref{Dimensionless}.\\

@@

Fig.~\ref{vd_vs_te}  shows either FIS or FIF, and a FIS+FIF pattern for desired
velocities $v_d>1.2\,$m/s. The  evacuation time attains a FIS+FIF pattern for
compression coefficients below  $k_n=1.2\,$E5. This means that ``soft''
individuals can attain this  behavioral pattern.  Notice that higher values of
$k_n$ allow  only FIS or FIF patterns. For the highest explored value
$k_n=1.2\,$E6, no  FIS can be seen at all. Besides, the evacuation pattern for
$k_n = 0$ and $k_n  = 2.6\,$E4 are very similar since the body force intensity
is of the same order or less than the social  force for these stiffness values.
\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the
pedestrian’s desired velocity. (b) Mean overlap as a function of the
pedestrians desired velocity. Each symbol indicates the $k_n$ value
corresponding to the body force (see the labels). The data corresponds to a
bottleneck with periodic boundary conditions (re-injecting pedestrians). The
average was taken over time and the pedestrians in the bottleneck. The sampling
was done every five seconds once the crowd reached the stationary state (say,
$t=20\,$s) until the end of the simulation ($t=1000\,$s). Color online only.}
\label{degree_overlap_vd}
\end{figure}

@@

We first computed the contact network in the same way  as in the bottleneck
geometry. Fig.~\ref{degree_dens} shows the  mean degree as a function of the
global density for different $k_n$ values (see  caption for details). The mean
degree vanishes at very low  densities because the pedestrians do not touch each
other. When the density  surpasses 4.5, a few pedestrians start to touch each
other,  raising the mean degree. As the density continuous increasing,  the mean
degree approaches the asymptotic value  of six. Degree six corresponds to the
maximum packing density for identical hard disks.\\

@@

\begin{figure}[!htbp]
\centering
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
\subfloat[]{\includegraphics[width=0.49\columnwidth]
{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the 
corridor at time $t=50\,$s. The global density was $\rho=6$.  The lines that 
connect the nodes (pedestrians) represent the contacts between them. The colors 
stand for the degree of the node (the number of pedestrians that are in contact 
with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary 
conditions and $v_d=1$. (a) corresponds to a simulation without body force and 
(b) corresponds to a simulation with $k_n=1.2$~E5. The friction coefficient and 
the other SFM parameters are the same as in Section \ref{bottleneck}. Color 
online only.}
\label{network_corridor}
\end{figure}

@@

Fig.~\ref{network_bottleneck} and  Fig.~\ref{network_corridor} illustrate the
connectivity differences between  the bottleneck and the corridor situation. The
bulk in  Fig.~\ref{network_bottleneck} appears more heavily connected among
``soft''  pedestrians than among ``stiff'' pedestrians (see both snapshots at
$v_d=10$).  The opposite occurs in Fig.~\ref{network_corridor}. This discrepancy
seems to  be related to the boundary conditions, since the same SFM parameters
were  applied on both situations. We may speculate that this phenomenon occurs
because, in the corridor, the lateral walls act like a  confining barrier that
forces the ``stiff'' pedestrian to increase his (her) contacts. On the contrary,
no real confining walls exist in the bottleneck situation (regardless the side
walls). \\

We can test our hypothesis by computing the mean overlap. 
Fig.~\ref{overlap_dens} shows this magnitude for the corridor situation. Notice 
that ``soft'' pedestrians attain more overlap than the ``stiff'' 
ones, as expected. This is in agreement with Fig.~\ref{overlap_vd} for the 
bottleneck situation (at a fixed value of $v_d$). The curves in 
Fig.~\ref{overlap_vd}, however, do not meet each other as in  
Fig.~\ref{overlap_dens} where all the curves meet each other at
high-density values. This phenomenon occurs due to space limitations
in the corridor that produces overlapping (almost) independently of 
the value of $k_n$. \\

@@

The above results represent an important step in the  investigation.
The inter-pedestrian connectivity differences between the bottleneck geometry
and the corridor geometry were not expected and opens two major questions:  how
do the pedestrians interact with the walls as a function of the body stiffness,
and consequently, how does  this affect the flux across the corridor.\\

@@

The $\langle v_x\rangle$ values in Fig.~\ref{kn_vs_vx_corridor}  were computed
at densities $\rho\geq 5.5$. According to  Fig.~\ref{overlap_dens}, at these
densities, the overlaps among pedestrians attain significant values.
Furthermore,  stiffness values $k_n\geq 10^4$ move the system to a more heavily
connected  stage, and thus, pedestrians have no choice but to walk at  (almost)
the same speed as his (her) neighbors. This  behavior may be envisaged as the
passage from a ``free''  walking movement to a constrained walking movement as
the stiffness increases. A more  physical picture would assimilate the former as
a ``fluid-like  state'' and the latter as a ``solid-like state''. When the
stiffness is  very high (say $k_n=$1.2~E6), all the pedestrians  are expected to
walk at a common velocity. The  pedestrians that walk in physical contact with
the wall are the ones who  determine the velocity of the whole crowd, as
discussed  below.\\

@@

From the results displayed above, we realize that the crowd behaves like a solid
for very stiff pedestrians. This means that the crowd can not be easily
``deformed''. In this context, deformation means that some parts of the crowd
may be allowed to move faster than other parts of the crowd. \\

@@

Where $\langle \cdot \rangle$ means the average taken over time. This definition
compares the velocity of the pedestrians close to the wall (boundary) with
respect to the velocity of the pedestrians at the center of the corridor
(center). Thus, the strain rate $\dot{\gamma}$ vanishes (no deformation) as the
stiffness level increases. This phenomenon is shown in
Fig.~\ref{strain_rate_vs_kn} where we can see that the strain rate drops for
high values of $k_n$. \\

We conclude this Section by stressing, once again, that stiffer pedestrians
attain opposite results in corridors with respect to bottlenecks. The walls in
the corridors play a critical role that prevents pedestrians from detaching from
each other. This effect can be observed as an increment in the connectivity
of the contact network and the flattening of the velocity profile (thus reducing
the strain rate). High enough stiffness values stuck the pedestrians
leading to a ``solidification'' of the crowd. All the pedestrians walk at almost
the same velocity at this stage. The pedestrians that are in contact with the
walls are the ones that determine the velocity of the whole crowd. We already
analyzed this situation in Ref~\cite{dorso_2019}. \\

@@

A further increment of $\mathcal{K}$ to $\mathcal{K}=$1371 (ten times the 
original SFM value) attains a plateau ($\rho>5$ p m$^{-2}$) before vanishing at 
very high densities. This occurs on either $\mathcal{K}_c=$ 0 and 
$\mathcal{K}_c = $ 68.\\

@@

Having adopted as fundamental philosophy of work that the parameters of the SFM
should be properly determined by requiring that the model reproduces
well established experimental data, we explored the effect on the pedestrians
dynamics of the (sometimes neglected) body force in the framework of the SFM. We
showed that the stiffness coefficient ($k_n$) has a significant impact on the
evacuation dynamics (bottleneck) and also in the dynamics of pedestrians walking
along a straight corridor.\\

@@

The authors want to thank the degree students Josefina Catoni and Ayelen Santos
 for providing data acquired at the subway in Buenos Aires, Argentina. \\
